Double White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification

### 2015-07-28 papers/wdmerger_I/paper.tex
Fix incorrect plot filename
Add reference to MacMillan's original calculation of the gravitational potential of a cube.

The only way to fully eliminate this effect is to use a test problem that
does not change with resolution. The obvious companion is a cube of
uniform density $\rho$, where now $R$ is half of the side length of
the cube. At each resolution we use the same $R$ as for the sphere,
which ensures that the cube always fills exactly the same fraction of
the domain and thus has the same mass, so the only improvement comes
from better sampling at higher resolution. The gravitational potential for this
object has been worked out analytically by \citet{waldvogel:1976} (see
also a similar result in \citealt{hummer:1996}). The potential is in
Equation 15 of that paper, though the last term is missing a factor of
$1/2$, which destroys the symmetry. Inserting this missing factor and
performing a simple coordinate transformation so that the center of
the cube is at the origin, the potential is
\begin{align}
  \Phi_{\text{cube}}(x,y,z) &= -G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
  &+ \left. y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\
  &\left. - \frac{x_i^2}{2}\,\text{tan}^{-1}\left(\frac{y_j z_k}{x_i r_{ijk}}\right) - \frac{y_j^2}{2}\,\text{tan}^{-1}\left(\frac{z_k x_i}{y_j r_{ijk}}\right) \right. \notag \\
  &- \left. \frac{z_k^2}{2}\,\text{tan}^{-1}\left(\frac{x_i y_j}{z_k r_{ijk}}\right)\right]
\end{align}
where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, $y_1 = R-y$, $z_0 =
R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
\texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent is
\texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
\texttt{atan2}). This formula is valid both inside and outside the
cube. The normalized $L^2$ error for this problem is also shown
in Figure \ref{fig:gravity_convergence}, and only for this problem 
do we obtain perfect second-order scaling at all resolutions.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.45]{plots/equal_energy_rot_0.eps}
  \caption{Absolute magnitude of the relative change in energy of two equal mass white dwarfs through 25 orbital periods,
           evolved in an inertial reference frame. The decline and recovery is a change in sign of the energy difference.
           \label{fig:energy_conservation_equal}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot_1.eps}
  \caption{Absolute magnitude of the relative change in angular momentum of two equal mass white dwarfs after 25 orbital periods, 
           evolved in a co-rotating reference frame. We consider only the component of the angular moment along the rotational 
           axis. The decline and recovery is a change in sign of the angular momentum difference.
           \label{fig:angular_momentum_conservation_unequal}}
\end{figure}

>>>>

The only way to fully eliminate this effect is to use a test problem that
does not change with resolution. The obvious companion is a cube of
uniform density $\rho$, where now $R$ is half of the side length of
the cube. At each resolution we use the same $R$ as for the sphere,
which ensures that the cube always fills exactly the same fraction of
the domain and thus has the same mass, so the only improvement comes
from better sampling at higher resolution. The gravitational potential for this
object has been worked out analytically by \citet{waldvogel:1976} (see
also a similar result by \citet{hummer:1996}, and an earlier calculation 
by \citet{macmillan:1958}). The potential is in
Equation 15 of that paper, though the last term is missing a factor of
$1/2$, which destroys the symmetry. Inserting this missing factor and
performing a simple coordinate transformation so that the center of
the cube is at the origin, the potential is
\begin{align}
  \Phi_{\text{cube}}(x,y,z) &= -G\rho\sum_{i,j,k=0}^1\left[x_i y_j\, \text{tanh}^{-1}\left(\frac{z_k}{r_{ijk}}\right)\right. \notag \\
  &+ \left. y_j z_k\, \text{tanh}^{-1}\left(\frac{x_i}{r_{ijk}}\right) + z_k x_i\, \text{tanh}^{-1}\left(\frac{y_j}{r_{ijk}}\right) \right.\notag \\
  &\left. - \frac{x_i^2}{2}\,\text{tan}^{-1}\left(\frac{y_j z_k}{x_i r_{ijk}}\right) - \frac{y_j^2}{2}\,\text{tan}^{-1}\left(\frac{z_k x_i}{y_j r_{ijk}}\right) \right. \notag \\
  &- \left. \frac{z_k^2}{2}\,\text{tan}^{-1}\left(\frac{x_i y_j}{z_k r_{ijk}}\right)\right]
\end{align}
where $x_0 = R + x$, $x_1 = R - x$, $y_0 = R + y$, $y_1 = R-y$, $z_0 =
R+z$, $z_1 = R-z$, and $r_{ijk} = \sqrt{x_i^2 + y_j^2 + z_k^2}$. In
\texttt{Fortran} and \texttt{C}, the inverse hyperbolic tangent is
\texttt{atanh} and the inverse tangent is \texttt{atan} (\textit{not}
\texttt{atan2}). This formula is valid both inside and outside the
cube. The normalized $L^2$ error for this problem is also shown
in Figure \ref{fig:gravity_convergence}, and only for this problem 
do we obtain perfect second-order scaling at all resolutions.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.45]{plots/equal_energy_rot0.eps}
  \caption{Absolute magnitude of the relative change in energy of two equal mass white dwarfs through 25 orbital periods,
           evolved in an inertial reference frame. The decline and recovery is a change in sign of the energy difference.
           \label{fig:energy_conservation_equal}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.45]{plots/unequal_angular_momentum_rot1.eps}
  \caption{Absolute magnitude of the relative change in angular momentum of two equal mass white dwarfs after 25 orbital periods, 
           evolved in a co-rotating reference frame. We consider only the component of the angular moment along the rotational 
           axis. The decline and recovery is a change in sign of the angular momentum difference.
           \label{fig:angular_momentum_conservation_unequal}}
\end{figure}
